Alex's Notes and Text Fragments for the JSS Manuscript

### 2021-08-06 paper/paper.Rmd
revised blocking section, reduced #figures

### Blocking

"Spatial Blocking" creates homogeneous blocks across the study area splitting observations into different zones [russ2010; @bahn2012; @wenger2012; @blockCV].
The creation of these zones can be diverse: one option is to supply rows and columns to subdivide the data based on a rectangular grid.
Others make use of landscape patterns such as crop zones [@russ2010] or species presence [@blockCV] to determine an initial partitioning.

Implementations of blocking methods often provide additional options to fine tune these initial partitioning algorithms.

\pkg{mlr3spatiotempcv} supports two "blocking" implementations: (1) "Spatial blocking" from package \pkg{blockCV} and (2) "Spatial Tiles" from \pkg{sperrorest}.

#### Spatial blocking --- blockCV

@@

The following subsections demonstrate the use of some of the introduced options.

##### Option: selection = "random"

@@

##### Option: selection = "systematic"

Here, aggregation follows the linear count from $1:n_{folds}$ from top to bottom.

@@

@@

This option ignores parameter `range` and always creates two partitions following a checkerboard pattern. The arguments `range` and `folds` are ignored in this case.

```{r block-checkerboard, fig.cap='Checkerboard resampling of square spatial blocks using the implementation in package blockCV (method "spcv\\_block" with option selection = "checkerboard" in mlr3spatiotempcv). The size of the squares is 1000 m, and eleven out of the 19 blocks were assigned to this test sample.', out.width="40%", fig.pos="ht!"}
resampling_block_checkerboard = rsmp("spcv_block",
  range = 1000, folds = 5,
  selection = "checkerboard")

@@

##### Rectangular blocks

Alternatively, arguments `rows` and `cols` can be specified instead of `range`.
This allows to create rectangular blocks which honor the supplied `rows` and `cols` settings.

@@

@@

### Clustering-based resampling

@@

autoplot(resampling_custom, size = 0.8, task = task) *
  ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.01)) *
  ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.01))

>>>>

### Spatial blocking

Spatial blocking divides the study area into different zones [russ2010; @bahn2012; @wenger2012; @blockCV].
Several methods have been implemented to create rectangular as well as irregularly shaped blocks. 
These blocks can be used for leave-one-block-out resampling, or they can themselves be grouped into $k$ partitions for a $k$-fold CV at the block level.

\pkg{mlr3spatiotempcv} supports multiple resampling methods that are based on blocks: 
(1) "Spatial blocking" from package \pkg{blockCV} implements $k$-fold CV resampling at the level of square or rectangular blocks, while (2) "Spatial tiles" from \pkg{sperrorest} offers leave-one-block-out resampling with square or rectangular, possibly rotated blocks.
In addition, irregularly shaped blocks are available (3) through coordinate-based clustering as a leave-one-block-out technique from \pkg{sperrorest} (method `"spcv_coords"`, Section&nbsp;\@ref(sec:cluster-cv)), and (4) as a CV at the block level where blocks are defined by a factor variable that defines zones or groups (method `"custom_cv"` in \pkg{mlr3}, Section&nbsp;\@ref(sec:custom-cv)).

#### Square and rectangular blocks --- blockCV

@@

This procedure supports both random and systematic resampling of square blocks (argument `selection = "random"` and `"systematic"`, respectively).
Block sizes in meters) are determined by the `range` argument.
Alternatively, rectangular blocks can be created by specifying the number of desired rows and columns (arguments `rows` and `cols`).

@@

In systematic resampling, the blocks are numbered row by row, and blocks $i+j\cdot\texttt{folds}$ are assigned to fold $i$.
This may create undesired patterns when the number of columns is equal to or a multiple of the number of folds.

@@

<!--

@@

-->

The *checkerboard* partitioning is a special case of a systematic block partitioning (`selection = "checkerboard"`).
It inherently supports only two folds, making it less appealing than the more commonly used five- or ten-fold resampling, which achieve larger training set sizes.

<!--
AB:
The text previously stated that the range argument is ignored. This can't be true, there must be some argument that controls the size of checkerboard fields.
-->

<!--
AB: Checkerboard pattern is just a special case of a systematic pattern and likely not of great practical relevance because there are only two folds -> omit figure.

```{r block-checkerboard, fig.cap='Checkerboard resampling of square spatial blocks using the implementation in package blockCV (method "spcv\\_block" with option selection = "checkerboard" in mlr3spatiotempcv). The size of the squares is 1000 m, and eleven out of the 19 blocks were assigned to this test sample.', out.width="40%", fig.pos="ht!"}
resampling_block_checkerboard = rsmp("spcv_block",
  range = 1000, folds = 2,
  selection = "checkerboard")

@@

-->

<!--
AB: This figure and subsubsubsection takes up a lot of space for very little additional information.

@@

-->

@@

### Clustering-based resampling {#sec:cluster-cv}

@@

# autoplot(resampling_custom, size = 0.8, task = task) *
#   ggplot2::scale_y_continuous(breaks = seq(-3.97, -4, -0.01)) *
#   ggplot2::scale_x_continuous(breaks = seq(-79.06, -79.08, -0.01))
